On electron acceleration by mildly-relativistic shocks: PIC simulations

### 2021-07-06 papers/Physica2021/Romansky_Physica2021.tex
turbulence 1

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.50\textwidth]{fig/Bnorm.png} 
		\caption{Temporal evolution of the transversal magnetic field.}
		\label{field}
	\end{figure}

	It is known, that development of Bell's instability can suppress electron acceleration \cite{Crumley2019,2021MNRAS.501.4837L,2021MNRAS.502.5065L}. We study this phenomena with different inclination angles of magnetic field. Spectrum of electrons is shown in figure \ref{spectrume}. One can see, that electron acceleration has strong dependence on inclination angle, and when it is more than critical value (at $\theta = 40^\circ$), non-thermal component almost disappears. For smaller angles electron distribution function at high energies $(\Gamma > 300)$ depends on time, and it maximizes at times about 1400 inverse proton plasma frequencies for angles $\theta=20^\circ, \theta=30^\circ$, and 2100 and inverse proton plasma frequencies for $\theta = 33^\circ$. After reaching the maximum value the electron distribution magnitude starts decreasing at high energies. It may by explained as consequence of magnetic field amplification. Instabilities increase the transverse magnetic field and also the angle between field and shock velocity. As a result shock becomes quasi-perpendicular and can not effectively accelerate particles \cite{Sironi2011,Romansky18}. On the other hand, proton distribution function constantly increases in time for sub-critical angles, and almost does not have non-thermal component for $\theta = 40^\circ$, as shown in figure \ref{spectrump}. We assume, that oscillating transverse field has the strongest influence on the smallest scales, and suppresses the electrons injection into acceleration process, while protons still can be accelerated. Also, one can see, that temporal evolutions of electron and proton spectrums are correlated. In case of inclination angle $\theta = 33^\circ$ proton distribution function grows slower then for smaller angles. And electron distribution in this case is maximizes at larger time. This correlation confirms the assumption, that instability suppressing electron acceleration is caused by accelerated protons.

	\begin{figure}[h!]
		\centering
		\begin{minipage}{0.48\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum20.png} 
		\end{minipage}\hfill
		\begin{minipage}{0.48\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum30.png} 
		\end{minipage}
		\begin{minipage}{0.48\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum33.png} 
		\end{minipage}
		\begin{minipage}{0.48\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum40.png} 
		\end{minipage}
		\caption{Temporal evolution of electron spectrum in shocks with different magnetic field inclination angles. Top row: left $\theta = 20^\circ$, right $\theta = 30^\circ$, bottom row: left $\theta = 33^\circ$, right $\theta = 40^\circ$}
		\label{spectrume}
	\end{figure}

\begin{figure}[h!]
	\centering
	\begin{minipage}{0.48\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump20.png} 
	\end{minipage}\hfill
	\begin{minipage}{0.48\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump30.png} 
	\end{minipage}
	\begin{minipage}{0.48\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump33.png} 
	\end{minipage}
	\begin{minipage}{0.48\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump40.png} 
	\end{minipage}
	\caption{Temporal evolution of proton spectrum in shocks with different magnetic field inclination angles. Top row: left $\theta = 20^\circ$, right $\theta = 30^\circ$, bottom row: left $\theta = 33^\circ$, right $\theta = 40^\circ$}
	\label{spectrump}
\end{figure}

@@

	Details of electron acceleration are important for correct interpretations of observations of the  synchrothron radiation from astrophysical sources. Particle-in-cell simulation has been used to demonstrate magnetic instabilities influence on electron acceleration in mildly-relativistic electron-ion shock. It is shown, that amplification of magnetic field suppresses formation of electron spectrum at high energies while proton spectrum can grow monotonously in time in subluminal shock. Also, we demonstrated that groth of proton spectrum is correlated with decreasing of electron acceleration, and resulting  electron spectrum strongly depends on magnetic field inclination angle. This effect should be taken into account for deriving parameters of source of synchrothron radiation via analysis of spectrum. Further particle-in-cell simulations with a wider spatial and energetical ranges will allow to study the influence of instabilities on electron spectrum more precisely.

>>>>

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.48\textwidth]{fig/Bnorm.png} 
		\caption{Temporal evolution of the transversal magnetic field.}
		\label{field}
	\end{figure}

	It is known, that development of Bell's instability can suppress electron acceleration \cite{Crumley2019,2021MNRAS.501.4837L,2021MNRAS.502.5065L}. We study this phenomena with different inclination angles of magnetic field. Spectrum of electrons is shown in figure \ref{spectrume}. One can see, that electron acceleration has strong dependence on inclination angle, and when it is more than critical value (at $\theta = 40^\circ$), non-thermal component almost disappears. For smaller angles electron distribution function at high energies $(\Gamma > 300)$ depends on time, and it maximizes at times about $1400$ inverse proton plasma frequencies $\omega_{pi}$ for angles $\theta=20^\circ, \theta=30^\circ$, and $2100$ and inverse proton plasma frequencies for $\theta = 33^\circ$. After reaching the maximum value the electron distribution magnitude starts decreasing at high energies. It may by explained as consequence of magnetic field amplification. Instabilities increase the transverse magnetic field and also the angle between field and shock velocity. As a result shock becomes quasi-perpendicular and can not effectively accelerate particles \cite{Sironi2011,Romansky18}. On the other hand, proton distribution function constantly increases in time for sub-critical angles, and almost does not have non-thermal component for $\theta = 40^\circ$, as shown in figure \ref{spectrump}. We assume, that oscillating transverse field has the strongest influence on the smallest scales, and suppresses the electrons injection into acceleration process, while protons still can be accelerated. Also, one can see, that temporal evolutions of electron and proton spectrums are correlated. In case of inclination angle $\theta = 33^\circ$ proton distribution function grows slower than for smaller angles, at time less than $2000 {\omega_{pi}}^{-1}$  it has very weak non-thermal component. And electron distribution in this case maximizes at larger time than for smaller angles. This correlation confirms the assumption, that instability suppressing electron acceleration is caused by accelerated protons.

	\begin{figure}[h!]
		\centering
		\begin{minipage}{0.47\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum20.png} 
		\end{minipage}\hfill
		\begin{minipage}{0.47\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum30.png} 
		\end{minipage}
		\begin{minipage}{0.47\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum33.png} 
		\end{minipage}
		\begin{minipage}{0.47\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum40.png} 
		\end{minipage}
		\caption{Temporal evolution of electron spectrum in shocks with different magnetic field inclination angles. Top row: left $\theta = 20^\circ$, right $\theta = 30^\circ$, bottom row: left $\theta = 33^\circ$, right $\theta = 40^\circ$}
		\label{spectrume}
	\end{figure}

\begin{figure}[h!]
	\centering
	\begin{minipage}{0.47\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump20.png} 
	\end{minipage}\hfill
	\begin{minipage}{0.47\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump30.png} 
	\end{minipage}
	\begin{minipage}{0.47\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump33.png} 
	\end{minipage}
	\begin{minipage}{0.47\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/spectrump40.png} 
	\end{minipage}
	\caption{Temporal evolution of proton spectrum in shocks with different magnetic field inclination angles. Top row: left $\theta = 20^\circ$, right $\theta = 30^\circ$, bottom row: left $\theta = 33^\circ$, right $\theta = 40^\circ$}
	\label{spectrump}
\end{figure}

@@

	Details of electron acceleration are important for correct interpretations of observations of the  synchrothron radiation from astrophysical sources. Particle-in-cell simulation has been used to demonstrate magnetic instabilities influence on electron acceleration in mildly-relativistic electron-ion shock. It is shown, that amplification of magnetic field suppresses formation of electron spectrum at high energies while proton spectrum can grow monotonously in time in subluminal shock. Also, we demonstrated that growth of proton spectrum is correlated with decreasing of electron acceleration, and resulting  electron spectrum strongly depends on magnetic field inclination angle. This effect should be taken into account for deriving parameters of source of synchrothron radiation via analysis of spectrum. Further particle-in-cell simulations with a wider spatial and energetical ranges will allow to study the influence of instabilities on electron spectrum more precisely.
